---
title: "PCA"
author: "Tin Buenafe"
date: "2023-08-07"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
                      warning = FALSE)
```

```{r preliminaries, include = FALSE}
# Define preliminaries
source("00_SetupGrid.R")
source("00_Preliminaries.R")
pacman::p_load(purrr, Hmisc, RColorBrewer, patchwork)
seasons <- c("jan-mar", "apr-jun", "jul-sept", "oct-dec")
PC <- c("PC1", "PC2")
figure_dir <- here::here(figure_dir, "PCA")

#### Prepare components for data frame ####
spp_list <- spec_dict %>% 
  dplyr::filter(!code %in% c("BON", "LIT")) %>% # remove bonitos and little tuna
  dplyr::select(code) %>% 
  pull() %>% 
  tolower()

df <- list() # empty list
for(i in 1:length(spp_list)) {
  obj <- readRDS(here::here(preds_dir, paste(toupper(spp_list[i]), "jan-mar.rds", sep = "_")))
  df[[i]] <- prepare_components(obj, spp_list[i])
}

df <- purrr::reduce(df, dplyr::left_join, by = c("cellID", "grid_100_category", "geometry"))
```
## Running PCA with covariances

For this I'm going to use __January-March__ as an example. First running the PCA with `cor = FALSE` to use covariances.

```{r PCA_cov, include = TRUE}
PCA <- stats::princomp(df %>% 
                         dplyr::select(-cellID, -grid_100_category, -geometry), cor = FALSE)
summary(PCA)
```

Then we use the PC scores to plot the map... for PC1

```{r PCA_cov_map1}
PC_scores <- PCA$scores[,1:2] %>% # first two axes
  tibble::as_tibble()

pc_obj <- df %>% 
  dplyr::select(cellID, grid_100_category, geometry) %>% 
  cbind(., PC_scores) %>% 
  sf::st_as_sf(crs = cCRS)

# Plot PC1
pc1 <- plotPC1_limits(pc_obj, "Comp.1", "PC1 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc1
```

and PC2

```{r PCA_cov_map2}
# Plot PC2
pc2 <- plotPC2_limits(pc_obj, "Comp.2", "PC2 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc2
```

We take the __loadings__ of each of the species (PC1) and plot them. Loadings are the association coefficients between the components and the variables. They are the covariances/correlations between the original variables and the unit-scaled components. They are directly comparable with association coefficients computed between the variables— covariances, correlations and other scalar products. See: [link1](https://stats.stackexchange.com/questions/2691/making-sense-of-principal-component-analysis-eigenvectors-eigenvalues/35653#35653) and [link2](https://stats.stackexchange.com/questions/143905/loadings-vs-eigenvectors-in-pca-when-to-use-one-or-another).

```{r PCA_cov_loadings}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(axis) {
  
  res <- PCA$loadings %>% 
    as.data.frame.matrix() %>% 
    tibble::rownames_to_column(var = "Species") %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1]  # get string of species
  
  res <- res[,2] %>%  # just get loadings
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] # rearrange the species
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
  
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Loadings",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("Comp.1")
gg <- plotCorrMat(res)
gg
```

And now we look at the __correlation coefficients__ of the Spearman correlation analysis between the PC scores and the individual species model outputs

```{r PCA_cov_spearman}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(season, axis) {
  
  res <- read_csv(here::here(pc_dir, paste("hotspots", season, "spearmancorr.csv", sep = "_"))) %>% 
    dplyr::rename(Species = `...1`) %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1] %>% 
    pull() # get string of species
  
  res <- res %>%  # just get loadings
    dplyr::filter(!Species %in% c("Comp.1", "Comp.2")) # remove axes

  res <- res[,2] %>% 
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] %>%  # rearrange the species
    tidyr::drop_na() # remove NAs
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Correlation coefficient",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("jan-mar", "Comp.1")
gg <- plotCorrMat(res)
gg
```

## Running PCA with correlation

First, I'm going to use __January-March__ as an example. PCA is ran with `cor = TRUE` to use correlation.

```{r PCA_cor, include = TRUE}
PCA <- stats::princomp(df %>% 
                         dplyr::select(-cellID, -grid_100_category, -geometry), cor = TRUE)
summary(PCA)
```

Then we use the PC scores to plot the map... for PC1

```{r PCA_cor_map1}
PC_scores <- PCA$scores[,1:2] %>% # first two axes
  tibble::as_tibble()

pc_obj <- df %>% 
  dplyr::select(cellID, grid_100_category, geometry) %>% 
  cbind(., PC_scores) %>% 
  sf::st_as_sf(crs = cCRS)

# Plot PC1
pc1 <- plotPC1_limits(pc_obj, "Comp.1", "PC1 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc1
```

and PC2

```{r PCA_cor_map2}
# Plot PC2
pc2 <- plotPC2_limits(pc_obj, "Comp.2", "PC2 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc2
```

And we take the __loadings__ of each of the species (PC1) and plot them...

```{r PCA_cor_loadings}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(axis) {
  
  res <- PCA$loadings %>% 
    as.data.frame.matrix() %>% 
    tibble::rownames_to_column(var = "Species") %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1]  # get string of species
  
  res <- res[,2] %>%  # just get loadings
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] # rearrange the species
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
  
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Loadings",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("Comp.1")
gg <- plotCorrMat(res)
gg
```

And now we look at the __correlation coefficients__ of the Spearman correlation analysis between the PC scores and the individual species model outputs

```{r PCA_cor_spearman}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(season, axis) {
  
  res <- read_csv(here::here(pc_dir, paste("hotspots", season, "spearmancorr.csv", sep = "_"))) %>% 
    dplyr::rename(Species = `...1`) %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1] %>% 
    pull() # get string of species
  
  res <- res %>%  # just get loadings
    dplyr::filter(!Species %in% c("Comp.1", "Comp.2")) # remove axes

  res <- res[,2] %>% 
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] %>%  # rearrange the species
    tidyr::drop_na() # remove NAs
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Correlation coefficient",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("jan-mar", "Comp.1")
gg <- plotCorrMat(res)
gg
```

## Then for the rest of the seasons

### April-June

```{r df_apr-jun, include = FALSE}
df <- list() # empty list
for(i in 1:length(spp_list)) {
  obj <- readRDS(here::here(preds_dir, paste(toupper(spp_list[i]), "apr-jun.rds", sep = "_")))
  df[[i]] <- prepare_components(obj, spp_list[i])
}

df <- purrr::reduce(df, dplyr::left_join, by = c("cellID", "grid_100_category", "geometry"))
```

```{r PCA2_cor, include = TRUE}
PCA <- stats::princomp(df %>% 
                         dplyr::select(-cellID, -grid_100_category, -geometry), cor = TRUE)
summary(PCA)
```

Then we use the PC scores to plot the map... for PC1

```{r PCA2_cor_map1}
PC_scores <- PCA$scores[,1:2] %>% # first two axes
  tibble::as_tibble()

pc_obj <- df %>% 
  dplyr::select(cellID, grid_100_category, geometry) %>% 
  cbind(., PC_scores) %>% 
  sf::st_as_sf(crs = cCRS)

# Plot PC1
pc1 <- plotPC1_limits(pc_obj, "Comp.1", "PC1 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc1
```

Loadings...

```{r PCA2_cor_loadings}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(axis) {
  
  res <- PCA$loadings %>% 
    as.data.frame.matrix() %>% 
    tibble::rownames_to_column(var = "Species") %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1]  # get string of species
  
  res <- res[,2] %>%  # just get loadings
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] # rearrange the species
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
  
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Loadings",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("Comp.1")
gg <- plotCorrMat(res)
gg
```

Spearman correlation coefficients...

```{r PCA2_cor_spearman}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(season, axis) {
  
  res <- read_csv(here::here(pc_dir, paste("hotspots", season, "spearmancorr.csv", sep = "_"))) %>% 
    dplyr::rename(Species = `...1`) %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1] %>% 
    pull() # get string of species
  
  res <- res %>%  # just get loadings
    dplyr::filter(!Species %in% c("Comp.1", "Comp.2")) # remove axes

  res <- res[,2] %>% 
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] %>%  # rearrange the species
    tidyr::drop_na() # remove NAs
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Correlation coefficient",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("apr-jun", "Comp.1")
gg <- plotCorrMat(res)
gg
```

### July-September

```{r df_jul-sept, include = FALSE}
df <- list() # empty list
for(i in 1:length(spp_list)) {
  obj <- readRDS(here::here(preds_dir, paste(toupper(spp_list[i]), "jul-sept.rds", sep = "_")))
  df[[i]] <- prepare_components(obj, spp_list[i])
}

df <- purrr::reduce(df, dplyr::left_join, by = c("cellID", "grid_100_category", "geometry"))
```

```{r PCA3_cor, include = TRUE}
PCA <- stats::princomp(df %>% 
                         dplyr::select(-cellID, -grid_100_category, -geometry), cor = TRUE)
summary(PCA)
```

Then we use the PC scores to plot the map... for PC1

```{r PCA3_cor_map1}
PC_scores <- PCA$scores[,1:2] %>% # first two axes
  tibble::as_tibble()

pc_obj <- df %>% 
  dplyr::select(cellID, grid_100_category, geometry) %>% 
  cbind(., PC_scores) %>% 
  sf::st_as_sf(crs = cCRS)

# Plot PC1
pc1 <- plotPC1_limits(pc_obj, "Comp.1", "PC1 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc1
```

Loadings...

```{r PCA3_cor_loadings}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(axis) {
  
  res <- PCA$loadings %>% 
    as.data.frame.matrix() %>% 
    tibble::rownames_to_column(var = "Species") %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1]  # get string of species
  
  res <- res[,2] %>%  # just get loadings
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] # rearrange the species
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
  
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Loadings",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("Comp.1")
gg <- plotCorrMat(res)
gg
```

Spearman correlation coefficients...

```{r PCA3_cor_spearman}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(season, axis) {
  
  res <- read_csv(here::here(pc_dir, paste("hotspots", season, "spearmancorr.csv", sep = "_"))) %>% 
    dplyr::rename(Species = `...1`) %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1] %>% 
    pull() # get string of species
  
  res <- res %>%  # just get loadings
    dplyr::filter(!Species %in% c("Comp.1", "Comp.2")) # remove axes

  res <- res[,2] %>% 
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] %>%  # rearrange the species
    tidyr::drop_na() # remove NAs
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Correlation coefficient",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("jul-sept", "Comp.1")
gg <- plotCorrMat(res)
gg
```

### October-December

```{r df_oct-dec, include = FALSE}
df <- list() # empty list
for(i in 1:length(spp_list)) {
  obj <- readRDS(here::here(preds_dir, paste(toupper(spp_list[i]), "oct-dec.rds", sep = "_")))
  df[[i]] <- prepare_components(obj, spp_list[i])
}

df <- purrr::reduce(df, dplyr::left_join, by = c("cellID", "grid_100_category", "geometry"))
```

```{r PCA4_cor, include = TRUE}
PCA <- stats::princomp(df %>% 
                         dplyr::select(-cellID, -grid_100_category, -geometry), cor = TRUE)
summary(PCA)
```

Then we use the PC scores to plot the map... for PC1

```{r PCA4_cor_map1}
PC_scores <- PCA$scores[,1:2] %>% # first two axes
  tibble::as_tibble()

pc_obj <- df %>% 
  dplyr::select(cellID, grid_100_category, geometry) %>% 
  cbind(., PC_scores) %>% 
  sf::st_as_sf(crs = cCRS)

# Plot PC1
pc1 <- plotPC1_limits(pc_obj, "Comp.1", "PC1 score") +
  theme(legend.text = element_text(size = 18, color = "black"))
pc1
```

Loadings...

```{r PCA4_cor_loadings}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(axis) {
  
  res <- PCA$loadings %>% 
    as.data.frame.matrix() %>% 
    tibble::rownames_to_column(var = "Species") %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1]  # get string of species
  
  res <- res[,2] %>%  # just get loadings
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] # rearrange the species
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
  
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Loadings",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("Comp.1")
gg <- plotCorrMat(res)
gg
```

Spearman correlation coefficients...

```{r PCA4_cor_spearman}
# DESCRIPTION: Prepare the object that would feed into plotting the correlation matrix

prepare_corrmat_obj <- function(season, axis) {
  
  res <- read_csv(here::here(pc_dir, paste("hotspots", season, "spearmancorr.csv", sep = "_"))) %>% 
    dplyr::rename(Species = `...1`) %>% 
    dplyr::select(Species, !!sym(axis)) %>% # just select species and principal component of interest
    dplyr::arrange(desc(!!sym(axis))) %>% # arrange in decreasing order of positive correlation
    dplyr::mutate(!!sym(axis) := as.numeric(!!sym(axis)))
  
  spp_str <- res[,1] %>% 
    pull() # get string of species
  
  res <- res %>%  # just get loadings
    dplyr::filter(!Species %in% c("Comp.1", "Comp.2")) # remove axes

  res <- res[,2] %>% 
    as.matrix()
  
  idx <- match(toupper(spp_str), spec_dict$code) # get indices of match
  spp_str_new <- spec_dict[idx,] %>%  # rearrange the species
    tidyr::drop_na() # remove NAs
  
  rownames(res) <- spp_str_new[,2] %>% 
    pull()
  
  return(res)
}

# DESCRIPTION: Plot correlation matrices
plotCorrMat <- function(x) {
  
  gg_cor <- ggcorrplot::ggcorrplot(x, 
                                   method = "circle",
                                   outline.color = "black"
  ) +
    expand_limits(y = c(0.5,1.5)) +
    scale_fill_gradient2(
      name = "Correlation coefficient",
      low = "#8c510a",
      mid = "#f5f5f5",
      high = "#01665e",
      limits = c(-1, 1)
      ) +
    scale_x_discrete(guide = guide_axis(angle = 90)) +
    theme_bw() +
    theme(legend.title = element_text(),
          legend.text = element_text(color = "black", size = 10),
          panel.grid = element_blank(),
          panel.grid.major = element_line(color = "grey86"),
          panel.border = element_blank(),
          axis.ticks = element_blank(),
          axis.text.y = element_blank(),
          axis.title = element_blank(),
          axis.text.x = element_text(color = "black", size = 12))
  
  return(gg_cor)
  
}

res <- prepare_corrmat_obj("oct-dec", "Comp.1")
gg <- plotCorrMat(res)
gg
```
