---
title: "Building sauries models"
author: "Tin Buenafe"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    css: "css.css"
---

```{r preliminaries, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, out.width = '100%', fig.align = 'center', fig.height = 3)
source("10a_SAU_Data.R") # Load SAU data
output_dir <- here::here("Output", "Models")
```

## Nishikawa plots

First, we plot the presence/absence data from the Nishikawa dataset for sauries.

### January-March

```{r nish_jm, echo = FALSE}
df <- `grid_SAU_jan-mar` %>% 
  dplyr::select(cellID, abundance, geometry) %>% 
  na.omit() %>% 
  dplyr::mutate(abundance_presence = ifelse(abundance > 0, yes = 1, no = 0)) %>% 
  sf::st_as_sf() %>% 
  dplyr::select(cellID, abundance_presence, everything()) # arrange

nish1 <- ggplot() +
  geom_sf(data = df, aes(fill = as.factor(abundance_presence), color = as.factor(abundance_presence))) +
  scale_discrete_manual(name = "Presence/Absence",
                        values = c(`1` = "#016450", `0` = "#a6bddb"),
                        aesthetics = c("fill", "color")) +
  geom_sf(data = landmass, fill = "black", color = "black", size = 1) +
  coord_sf(xlim = st_bbox(grid)$xlim, ylim = st_bbox(grid)$ylim) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        panel.border = element_blank(),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
nish1
```

### April-June

```{r nish_aj, echo = FALSE}
df <- `grid_SAU_apr-jun` %>% 
  dplyr::select(cellID, abundance, geometry) %>% 
  na.omit() %>% 
  dplyr::mutate(abundance_presence = ifelse(abundance > 0, yes = 1, no = 0)) %>% 
  sf::st_as_sf() %>% 
  dplyr::select(cellID, abundance_presence, everything()) # arrange

nish2 <- ggplot() +
  geom_sf(data = df, aes(fill = as.factor(abundance_presence), color = as.factor(abundance_presence))) +
  scale_discrete_manual(name = "Presence/Absence",
                        values = c(`1` = "#016450", `0` = "#a6bddb"),
                        aesthetics = c("fill", "color")) +
  geom_sf(data = landmass, fill = "black", color = "black", size = 1) +
  coord_sf(xlim = st_bbox(grid)$xlim, ylim = st_bbox(grid)$ylim) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        panel.border = element_blank(),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
nish2
```

### July-September

```{r nish_js, echo = FALSE}
df <- `grid_SAU_jul-sept` %>%
  dplyr::select(cellID, abundance, geometry) %>%
  na.omit() %>%
  dplyr::mutate(abundance_presence = ifelse(abundance > 0, yes = 1, no = 0)) %>%
  sf::st_as_sf() %>%
  dplyr::select(cellID, abundance_presence, everything()) # arrange

nish3 <- ggplot() +
  geom_sf(data = df, aes(fill = as.factor(abundance_presence), color = as.factor(abundance_presence))) +
  scale_discrete_manual(name = "Presence/Absence",
                        values = c(`1` = "#016450", `0` = "#a6bddb"),
                        aesthetics = c("fill", "color")) +
  geom_sf(data = landmass, fill = "black", color = "black", size = 1) +
  coord_sf(xlim = st_bbox(grid)$xlim, ylim = st_bbox(grid)$ylim) +
  theme_bw() +
  theme(legend.position = "bottom",
          axis.title = element_blank(),
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 10),
          panel.border = element_blank(),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
nish3
```

### October-December

```{r nish_od, echo = FALSE}
df <- `grid_SAU_oct-dec` %>%
  dplyr::select(cellID, abundance, geometry) %>%
  na.omit() %>%
  dplyr::mutate(abundance_presence = ifelse(abundance > 0, yes = 1, no = 0)) %>%
  sf::st_as_sf() %>%
  dplyr::select(cellID, abundance_presence, everything()) # arrange

nish4 <- ggplot() +
  geom_sf(data = df, aes(fill = as.factor(abundance_presence), color = as.factor(abundance_presence))) +
  scale_discrete_manual(name = "Presence/Absence",
                        values = c(`1` = "#016450", `0` = "#a6bddb"),
                        aesthetics = c("fill", "color")) +
  geom_sf(data = landmass, fill = "black", color = "black", size = 1) +
  coord_sf(xlim = st_bbox(grid)$xlim, ylim = st_bbox(grid)$ylim) +
  theme_bw() +
  theme(legend.position = "bottom",
          axis.title = element_blank(),
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 10),
          panel.border = element_blank(),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
nish4
```

## Base model (Model 1)

Then, we build the basic sauries model with all the predictors included. Note that the environmental predictors are mean values over 1956-1981.

1. longitude
2. latitude
3. season
4. temperature ($^{\circ} C$)
5. oxygen ($mol \; m^{-3}$)
6. pH
7. chlorophyll ($mol \; m^{-3}$)
8. salinity (ppt)
9. mixed layer thickness (m)
10. nitrate ($mol \; m^{-3}$)
11. phosphate ($mol \; m^{-3}$)
12. ammonium ($mol \; m^{-3}$)
13. broad-scale thermal gradient ($\Delta  ^{\circ} C \; km^{-1}$)
14. broad-scale salinity gradient ($ppt \; km^{-1}$)
15. eddy kinetic energy ($m^{-2} \; s^{-2}$)
16. depth (m)
17. distance from nearest coast (km)
18. mean probability of adult occurrence for three sauries species (*Cololabis saira*, *Cololabis adocetus*, and *Scomberesox saurus*)

```{r model1, echo = FALSE}
SAU_model1 <- readRDS("Output/Models/SAU_model1.rds")

print(paste0("training AUC: ", SAU_model1$self.statistics$discrimination)) # Training AUC Score

# Predict to the testing dataset
preds <- gbm::predict.gbm(SAU_model1, test, n.trees = SAU_model1$gbm.call$best.trees, type = "response")
print(paste0("testing AUC: ", get_testAUC(test$abundance_presence, preds))) # Print testing AUC

# Plot maps
train_tmp <- train %>%
  dplyr::mutate(model = SAU_model1$fitted)
test_tmp <- test %>%
  dplyr::mutate(model = preds)
```

Then, we extrapolate for the rest of $40^{\circ}N$-$40^{\circ}S$ and present seasonal distribution maps. The distribution maps are shown side-by-side with the Nishikawa maps.

### January-March
```{r model1_jm, echo = FALSE}
# January-March
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "jan-mar", # season
                        SAU_predict_season1, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_jan-mar` # grid of species for specific season
)

gg1 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg1 + nish1
```

And, unlike yellowfin tuna, the adult sauries have a more restricted range. We show the plot with hatches in areas that have adult probabilities < 0.01.

```{r hatched1_jm, echo = FALSE}
hatch1 <- plotHatch(gg1, gg, SAU_ds1)
hatch1
```

### April-June
```{r model1_aj, echo = FALSE}
# April-June
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "apr-jun", # season
                        SAU_predict_season2, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_apr-jun` # grid of species for specific season
)

gg2 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg2 + nish2
```

```{r hatched1_aj, echo = FALSE}
hatch2 <- plotHatch(gg2, gg, SAU_ds2)
hatch2
```

### July-September
```{r model1_js, echo = FALSE}
# July-September
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "jul-sept", # season
                        SAU_predict_season3, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_jul-sept` # grid of species for specific season
)

gg3 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg3 + nish3
```

```{r hatched1_js, echo = FALSE}
hatch3 <- plotHatch(gg3, gg, SAU_ds3)
hatch3
```

### October-December
```{r model1_od, echo = FALSE}
# October-December
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "oct-dec", # season
                        SAU_predict_season4, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_oct-dec` # grid of species for specific season
)

gg4 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg4 + nish4
```

```{r hatched1_od, echo = FALSE}
hatch4 <- plotHatch(gg4, gg, SAU_ds4)
hatch4
```

## Model without geographical location (Model 2)

```{r model2, echo = FALSE}
SAU_model2 <- readRDS("Output/Models/SAU_model2.rds")

print(paste0("training AUC: ", SAU_model2$self.statistics$discrimination)) # Training AUC Score

# Predict to the testing dataset
preds <- gbm::predict.gbm(SAU_model2, test, n.trees = SAU_model2$gbm.call$best.trees, type = "response")
print(paste0("testing AUC: ", get_testAUC(test$abundance_presence, preds))) # Print testing AUC

# Plot maps
train_tmp <- train %>%
  dplyr::mutate(model = SAU_model2$fitted)
test_tmp <- test %>%
  dplyr::mutate(model = preds)
```

Again, each seasonal distribution map is shown side-by-side with its corresponding Nishikawa seasonal chart.

### January-March
```{r model2_jm, echo = FALSE}
# January-March
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "jan-mar", # season
                        SAU_predict_season1, # rest of the ocean cells
                        SAU_model2, # BRT model
                        `grid_SAU_jan-mar` # grid of species for specific season
)

gg1 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg1 + nish1
```

```{r hatched2_jm, echo = FALSE}
hatch1 <- plotHatch(gg1, gg, SAU_ds1)
hatch1
```

### April-June
```{r model2_aj, echo = FALSE}
# April-June
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "apr-jun", # season
                        SAU_predict_season2, # rest of the ocean cells
                        SAU_model2, # BRT model
                        `grid_SAU_apr-jun` # grid of species for specific season
)

gg2 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg2 + nish2
```

```{r hatched2_aj, echo = FALSE}
hatch2 <- plotHatch(gg2, gg, SAU_ds2)
hatch2
```

### July-September

```{r model2_js, echo = FALSE}
# July-September
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "jul-sept", # season
                        SAU_predict_season3, # rest of the ocean cells
                        SAU_model2, # BRT model
                        `grid_SAU_jul-sept` # grid of species for specific season
)

gg3 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg3 + nish3
```

```{r hatched2_js, echo = FALSE}
hatch3 <- plotHatch(gg3, gg, SAU_ds3)
hatch3
```

### October-December
```{r model2_od, echo = FALSE}
# October-December
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "oct-dec", # season
                        SAU_predict_season4, # rest of the ocean cells
                        SAU_model2, # BRT model
                        `grid_SAU_oct-dec` # grid of species for specific season
)

gg4 <- plotModel(gg) +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm")) # Plot the squished model
gg4 + nish4
```

```{r hatched2_od, echo = FALSE}
hatch4 <- plotHatch(gg4, gg, SAU_ds4)
hatch4
```

## Increasing confidence of distribution maps.

For this section, we use Model 1 (full model). We use the same $10 \times 10$ grid.

``` {r grid10x10, echo = FALSE}
# Objects for plotting seasonal maps using model 1
preds <- gbm::predict.gbm(SAU_model1, test, n.trees = SAU_model1$gbm.call$best.trees, type = "response")
train_tmp <- train %>%
  dplyr::mutate(model = SAU_model1$fitted)
test_tmp <- test %>%
  dplyr::mutate(model = preds)

# Build the 10x10 grid
grid_100 <- sf::st_make_grid(Bndry,
                         square = TRUE,
                         cellsize = c(10,10),
                         what = "polygons") %>%
  sf::st_sf()

# First get all the PUs partially/wholly within the planning region
logi_Reg <- sf::st_centroid(grid_100) %>%
  sf::st_intersects(Bndry) %>%
  lengths > 0 # Get logical vector instead of sparse geometry binary

grid_100 <- grid_100[logi_Reg, ] # Get TRUE

grid_100 %<>%
  dplyr::mutate(cellID = dplyr::row_number()) # Add a cell ID reference
```

### January-March

Similar to the yellowfin tuna, for each season, we associate the $10 \times 10$ grid with the $1 \times 1$ grid cells. Then, we limit the area to $10 \times 10$ grid cells with sampling points.

``` {r assoc_jm, echo = FALSE}
# Associate seasonal grids with the 10x10 grid
full_grid <- associateGrids(`grid_SAU_jan-mar`, grid_100)

# Create January-March seasonal map using model 1
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "jan-mar", # season
                        SAU_predict_season1, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_jan-mar` # grid of species for specific season
)

# Filter model outputs to those 10x10 grid cells with sampling points within them
gg_new <- restrictArea(gg, full_grid)

plotModel(gg_new) +
  theme(plot.margin = unit(c(0,0.5,0,0.5), "cm"))
```


We only leave $10 \times 10$ grid cells that have sampling points within a certain % area threshold. We first do this for a more conservative __25%__ threshold.

```{r thresh25_jm, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 25% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             25)

gg_25 <- plotModel(gg_filt) +
  ggtitle("January-March: 25% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_25 + nish1
```

Let's also visualize that side-by-side with the hatched plot that's filtered to just the remaining $10 \times 10$ grid cells.

```{r threshatched25_jm, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds1 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt1 <- plotHatch(gg_25, gg_filt, df_filt)
gg_25 + hatch_filt1
```

We now put hatches over areas of lower confidence.
``` {r hatchFull25_jm, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

We also do it for a more liberal __10%__ threshold.
```{r thresh10_jm, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 10% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             10)

gg_10 <- plotModel(gg_filt) +
  ggtitle("January-March: 10% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_10 + nish1
```

```{r threshatched10_jm, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds1 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt1 <- plotHatch(gg_10, gg_filt, df_filt)
gg_10 + hatch_filt1
```

``` {r hatchFull10_jm, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

Then, we replicate this across the 3 other seasons...

### April-June

``` {r assoc_aj, echo = FALSE}
# Associate seasonal grids with the 10x10 grid
full_grid <- associateGrids(`grid_SAU_apr-jun`, grid_100)

# Create April-June seasonal map using model 1
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "apr-jun", # season
                        SAU_predict_season2, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_apr-jun` # grid of species for specific season
)

# Filter model outputs to those 10x10 grid cells with sampling points within them
gg_new <- restrictArea(gg, full_grid)

plotModel(gg_new) +
  theme(plot.margin = unit(c(0,0.5,0,0.5), "cm"))
```

```{r thresh25_aj, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 25% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             25)

gg_25 <- plotModel(gg_filt) +
  ggtitle("April-June: 25% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_25 + nish2
```

```{r threshatched25_aj, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds2 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt2 <- plotHatch(gg_25, gg_filt, df_filt)
gg_25 + hatch_filt2
```

``` {r hatchFull25_aj, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

```{r thresh10_aj, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 10% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             10)

gg_10 <- plotModel(gg_filt) +
  ggtitle("April-June: 10% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_10 + nish2
```

```{r threshatched10_aj, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds2 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt2 <- plotHatch(gg_10, gg_filt, df_filt)
gg_10 + hatch_filt2
```

``` {r hatchFull10_aj, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

### July-September

``` {r assoc_js, echo = FALSE}
# Associate seasonal grids with the 10x10 grid
full_grid <- associateGrids(`grid_SAU_jul-sept`, grid_100)

# Create July-September seasonal map using model 1
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "jul-sept", # season
                        SAU_predict_season3, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_jul-sept` # grid of species for specific season
)

# Filter model outputs to those 10x10 grid cells with sampling points within them
gg_new <- restrictArea(gg, full_grid)

plotModel(gg_new) +
  theme(plot.margin = unit(c(0,0.5,0,0.5), "cm"))
```

```{r thresh25_js, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 25% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             25)

gg_25 <- plotModel(gg_filt) +
  ggtitle("July-September: 25% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_25 + nish3
```

```{r threshatched25_js, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds3 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt3 <- plotHatch(gg_25, gg_filt, df_filt)
gg_25 + hatch_filt3
```
``` {r hatchFull25_js, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

```{r thresh10_js, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 10% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             10)

gg_10 <- plotModel(gg_filt) +
  ggtitle("July-September: 10% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_10 + nish3
```

```{r threshatched10_js, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds3 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt3 <- plotHatch(gg_10, gg_filt, df_filt)
gg_10 + hatch_filt3
```

``` {r hatchFull10_js, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

### October-December

``` {r assoc_od, echo = FALSE}
# Associate seasonal grids with the 10x10 grid
full_grid <- associateGrids(`grid_SAU_oct-dec`, grid_100)

# Create October-December seasonal map using model 1
gg <- create_speciesMap(train_tmp, # training object with model column (fitted values)
                        test_tmp, # testing object with model column (predictions)
                        "oct-dec", # season
                        SAU_predict_season4, # rest of the ocean cells
                        SAU_model1, # BRT model
                        `grid_SAU_oct-dec` # grid of species for specific season
)

# Filter model outputs to those 10x10 grid cells with sampling points within them
gg_new <- restrictArea(gg, full_grid)

plotModel(gg_new) +
  theme(plot.margin = unit(c(0,0.5,0,0.5), "cm"))
```

```{r thresh25_od, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 25% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             25)

gg_25 <- plotModel(gg_filt) +
  ggtitle("October-December: 25% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_25 + nish4
```

```{r threshatched25_od, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds4 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt4 <- plotHatch(gg_25, gg_filt, df_filt)
gg_25 + hatch_filt4
```

``` {r hatchFull25_od, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```

```{r thresh10_od, echo = FALSE}
# Restrict model outputs to just 10x10 grid cells that have at least 10% of its area being sampling points.
gg_filt <- restrictThreshold(full_grid,
                             gg,
                             10)

gg_10 <- plotModel(gg_filt) +
  ggtitle("October-December: 10% threshold") +
  theme(legend.text = element_text(size = 8),
        legend.title = element_text(size = 10),
        plot.margin = unit(c(0,0.5,0,0.5), "cm"))
gg_10 + nish4
```

```{r threshatched10_od, echo = FALSE}
# Filter original grid to only these cells for plotting
df_filt <- SAU_ds4 %>%
  dplyr::filter(cellID %in% gg_filt$cellID)

hatch_filt4 <- plotHatch(gg_10, gg_filt, df_filt)
gg_10 + hatch_filt4
```

``` {r hatchFull10_od, echo = FALSE}
plotConfidence(gg_filt, full_grid, gg)
```